---
title: "p8105_hw3_xw2758"
author: "Xinyi Wang"
date: "10/10/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```

Let's make a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Let's make a table!!

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```

Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```


## Problem 2

### Fisrt, I will load, tidy, and otherwise wrangle the data.

```{r echo=TRUE}
acc_df = read_csv("~/R_columbia/p8105/p8105_hw3_xw2758/data/accel_data.csv") %>% 
 janitor::clean_names() %>% 
 pivot_longer(
  activity_1:activity_1440,
  names_to = "minute",
  names_prefix = "activity.",
  values_to = "activity_counts"
 ) %>% 
 mutate(
  weekday_vs_weekend = case_when(
   day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday",
   day  %in% c("Saturday","Sunday") ~ "weekend"
  )
 ) %>% 
  mutate(
    week = as.factor(week),
    day_id = as.factor(day_id), 
    day = as.factor(day),
    minute = as.numeric(minute),
    activity_counts = as.numeric(activity_counts),
    weekday_vs_weekend = as.factor(weekday_vs_weekend)
  )  %>% 
 mutate(day = forcats::fct_relevel(day,
                                   c("Monday","Tuesday","Wednesday","Thursday","Friday", "Saturday","Sunday"))
         )
```

acc_df dataset has `r nrow(acc_df)` observations and `r ncol(acc_df)`variables, including `r names(acc_df)` .

### Aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent

```{r echo=TRUE}
acc_df %>% 
 group_by(week,day) %>% 
  summarise(total_activity = sum(activity_counts)) %>% 
  pivot_wider(
    names_from = "day", 
    values_from = "total_activity") %>% 
 knitr::kable()
```

The acticivity counts are quite steady without too much change. However, the Saturday and Sunday in week 4&5 are quite low.

### Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
acc_df %>% 
  group_by(day_id, day, week) %>% 
  ggplot(aes(x = minute, y = activity_counts, color = day, group = day_id)) + 
  geom_point(alpha = 0.01) +
  theme_minimal()+
  theme(legend.position = "bottom") +
  geom_smooth(se = F, method = 'loess') + 
  labs(title = "Activities plot",
       x="Time(mintute)",
       y= "activity count",
       caption = "24-hour activity acceleremoter data of 63 year-old male with BMI 25") + 
  scale_x_continuous(
    breaks = c(0,300,600,900,1200,1500), 
    labels = c("0", "300", "600", "900", "1200", "1500"),
    limits = c(0, 1500)) + 
  scale_y_continuous(
    trans = "log",
    breaks = c(0,2000,4000,6000,8000,10000), 
    labels = c("0", "2000", "4000", "6000", "8000","10000"),
    limits = c(0, 8000))
```







